---
title: "EDA Real Estate"
output:
  pdf_document: default
  html_notebook: default
---

```{r}
# Load Data
library(xlsx)
listing = read.xlsx("HES_TEMPLATES_v2.xlsx", sheetIndex = 1)
economy = read.xlsx("HES_TEMPLATES_v2.xlsx", sheetIndex = 2, startRow = 1, endRow = 382)
```


```{r}
# column names for economy dataset
colnames(economy)
```

```{r}
# column names for listing dataset
colnames(listing)
```
```{r}
head(listing)
```




```{r}
# Fit a linear regression model on all variables in listing
fit_all <- lm(Avg..Median.Listing.Price.Yy ~ Avg..Active.Listing.Count.Yy + Avg..Ldpviews.Per.Property.Yy + Avg..Median.Listing.Price + Avg..Active.Listing.Count + Avg..Ldpviews.Per.Property + Avg..Median.Dom + Avg.ActiveListingsper1000HH + factor(Region), data=listing)
summary(fit_all)
```

Only Avg..Active.Listing.Count.Yy is statistically significant in top 50, therefore using `listings`.


```{r}
# compute VIF, and remove variables if VIF > 5 since they have a problem estimating
library(car)
vif(fit_all)
```

From above, we can see that all the values do not have a problem estimating the coefficients, hence we can keep the existing variables. 


```{r}
# Trying Stepwise Regression with all variables
source("http://people.fas.harvard.edu/~mparzen/stat100/model_select.txt")
model.select(fit_all, verbose = FALSE)
```

Stepwise Regression results in Avg..Active.Listing.Count.Yy in using top 50

in using all
Avg..Median.Listing.Price.Yy ~ Avg..Active.Listing.Count.Yy + Avg..Ldpviews.Per.Property.Yy + Avg..Median.Listing.Price +  Avg..Median.Dom + factor(Region) 

lm(formula = Avg..Median.Listing.Price.Yy ~ Avg..Active.Listing.Count.Yy + 
    Avg..Ldpviews.Per.Property.Yy + Avg..Median.Listing.Price + 
    Avg..Median.Dom + factor(Region)



```{r}
# fit the model with the stepwise results
fit_stepwise_res <- lm(formula = Avg..Median.Listing.Price.Yy ~ Avg..Active.Listing.Count.Yy + Avg..Ldpviews.Per.Property.Yy + Avg..Median.Listing.Price + Avg..Median.Dom + factor(Region), data=listing)
summary(fit_stepwise_res)
```


```{r}
# just confirming stepwise results here
model.select(fit_stepwise_res, verbose = FALSE)
```


```{r}
# merge stepwise results with economy dataset
vars <- c("Avg..Median.Listing.Price.Yy", "Avg..Active.Listing.Count.Yy", "Avg..Ldpviews.Per.Property.Yy", "Avg..Median.Listing.Price", "Avg..Median.Dom", "RankHH")
stepwise_listing_df <- listing[vars]
listing_economy <- merge(stepwise_listing_df, economy, by="RankHH")
colnames(listing_economy)
```


```{r}
# run merged results
merge_stepwise_fit <- lm(Avg..Median.Listing.Price.Yy ~ Avg..Median.Listing.Price.Yy + + Avg..Active.Listing.Count.Yy + Avg..Ldpviews.Per.Property.Yy + Avg..Median.Listing.Price + Avg..Median.Dom + factor(Region) + Sale.Px.Recovery + Sale.Price.Yoy, data=listing_economy)
summary(merge_stepwise_fit)
```

```{r, warning = FALSE}
model.select(merge_stepwise_fit, verbose = FALSE)
```



Running Alternative Tests

```{r}
# Load Data
library(xlsx)
listing = read.xlsx("HES_TEMPLATES_v2.xlsx", sheetIndex = 1)
economy = read.xlsx("HES_TEMPLATES_v2.xlsx", sheetIndex = 2, startRow = 1, endRow = 382)

listing_economy = merge(listing, economy, by="RankHH")
colnames(listing_economy)
```


```{r}
boxplot(listing_economy$Avg..Median.Listing.Price.Yy, main="Avg..Median.Listing.Price.Yy")
boxplot(listing_economy$Avg..Active.Listing.Count.Yy, main="Avg..Active.Listing.Count.Yy")
boxplot(listing_economy$Avg..Ldpviews.Per.Property.Yy, main="Avg..Ldpviews.Per.Property.Yy")
boxplot(listing_economy$Avg..Median.Dom.Yy, main="Avg..Median.Dom.Yy")

```


