---
title: "RE Project Notes RS"
author: "Ravi Sharma"
date: "April 3, 2019"
output:
  html_document:
    df_print: paged
---

Load package for working with Excel files
```{r}
library(xlsx)
```

Load listing data
```{r}
listing <- read.xlsx("HES_TEMPLATES_v2.xlsx", 1)

dim(listing)
```

Load economy data
```{r}
economy <- read.xlsx("HES_TEMPLATES_v2.xlsx", sheetIndex = 2, startRow = 1, endRow = 382)

dim(economy)

```

Header records:
```{r}
head(listing)
```

```{r}
tail(economy)
```

## As discussed in the group (...add details), to consider a subset comprising of top 100 ranking city areas, a smaller dataset is created:
```{r}
listing100 <- data.frame(listing[1:100,])
dim(listing100)
```


## Summary of variables in Listings (top 100 ranks subset of data):
### To begin with, list all variables:
```{r}
names(listing)
#summary(listing)
```
### 04. Regions represented in the data:
Summary table:
```{r}
summary(listing100$Region)
```
Visual:
```{r}
barplot(table(listing100$Region), main = "Regions in top 100 listing centers", xlab="Regions")
```

Comments:


### 05. Median.Listing.Price.Yy change:
Summary table:
```{r}
summary(listing100$Avg..Median.Listing.Price.Yy)
```
Visual:
```{r}
boxplot(listing100$Avg..Median.Listing.Price.Yy)
```


### 06. Active.Listing.Count.Yy change:
Summary table:
```{r}
summary(listing100$Avg..Active.Listing.Count.Yy)
```
Visual:
```{r}
boxplot(listing100$Avg..Active.Listing.Count.Yy)
```

### 07. Ldpviews.Per.Property.Yy change:
Summary table:
```{r}
summary(listing100$Avg..Ldpviews.Per.Property.Yy)
```
Visual:
```{r}
boxplot(listing100$Avg..Ldpviews.Per.Property.Yy)
```


### 08. Median.Dom.Yy change:
Summary table:
```{r}
summary(listing100$Avg..Median.Dom.Yy)
```
Visual:
```{r}
boxplot(listing100$Avg..Median.Dom.Yy)
```

### 09. Median.Listing.Price :
Summary table:
```{r}
summary(listing100$Avg..Median.Listing.Price)
```
Visual:
```{r}
boxplot(listing100$Avg..Median.Listing.Price)
```


### 10. Active.Listing.Count :
Summary table:
```{r}
summary(listing100$Avg..Active.Listing.Count)
```
Visual:
```{r}
boxplot(listing100$Avg..Active.Listing.Count)
```


### 11. Ldpviews.Per.Property :
Summary table:
```{r}
summary(listing100$Avg..Ldpviews.Per.Property)
```
Visual:
```{r}
boxplot(listing100$Avg..Ldpviews.Per.Property)
```


### 12. Median.Dom :
Summary table:
```{r}
summary(listing100$Avg..Median.Dom)
```
Visual:
```{r}
boxplot(listing100$Avg..Median.Dom)
```


### 13. ActiveListingsper1000HH :
Summary table:
```{r}
summary(listing100$Avg.ActiveListingsper1000HH)
```
Visual:
```{r}
boxplot(listing100$Avg.ActiveListingsper1000HH)
```




## Try regression with everything in:
```{r}
fit1 <- lm(Avg..Median.Listing.Price.Yy ~ Avg..Active.Listing.Count.Yy + Avg..Ldpviews.Per.Property.Yy + Avg..Median.Dom.Yy + Avg..Median.Listing.Price + Avg..Active.Listing.Count + Avg..Ldpviews.Per.Property + Avg..Median.Dom + Avg.ActiveListingsper1000HH , data=listing100)

summary(fit1)
```

### load script for stepwise regression:
```{r}
source("http://people.fas.harvard.edu/~mparzen/stat100/model_select.txt")
```

### Now run stepwise regression:
```{r}
model.select(fit1, verbose = FALSE)
```


## Let us also include the categorical variable region to list of predictors:
```{r}
fit2 <- lm(Avg..Median.Listing.Price.Yy ~ Avg..Active.Listing.Count.Yy + Avg..Ldpviews.Per.Property.Yy + Avg..Median.Dom.Yy + Avg..Median.Listing.Price + Avg..Active.Listing.Count + Avg..Ldpviews.Per.Property + Avg..Median.Dom + Avg.ActiveListingsper1000HH + factor(Region) , data=listing100)

summary(fit2)
```

### Now run stepwise regression again:
```{r}
model.select(fit2, verbose = FALSE)
```


## Let us look at one-to-one relationship between Y and X and between Xs:
```{r}
pairs(listing100[,5:13])
```

## Now let us look at the suggested predictors individually, in detail:

### 
```{r}
fit3 <- lm(Avg..Median.Listing.Price.Yy ~ Avg..Active.Listing.Count.Yy, data=listing100)

summary(fit3)
```
Scatterplot and regression line
```{r}
plot(listing100$Avg..Active.Listing.Count.Yy, listing100$Avg..Median.Listing.Price.Yy)
abline(fit3, col="orange", lwd=3)
```

The plot is somewhat convincing. Let us look for influential points
```{r}
plot(fit3,4)
```

Let us list the two longest on Cook's distance:
```{r}
listing100[c(31,89),1:4]
```

Let us remove those two:
```{r}
listing98<-listing100[!(listing100$RankHH == 31 | listing100$RankHH == 89),]
dim(listing98)
```

And now run the regression again:
```{r}
fit4 <- lm(Avg..Median.Listing.Price.Yy ~ Avg..Active.Listing.Count.Yy, data=listing98)

summary(fit4)
```
Slightly improved Adjusted R-squared, residual standard error reduces slightly. Let us see the plots again:
```{r}
plot(listing98$Avg..Active.Listing.Count.Yy, listing98$Avg..Median.Listing.Price.Yy)
abline(fit3, col="orange", lwd=3)
```



### Let us get economy data into the mix:
## First let us create the combined dataset:
```{r}
listing_eco_98 <- merge(listing98, economy, by="RankHH")
dim(listing_eco_98)
```

## we now have a long list of columns. Here is the list:
```{r}
names(listing_eco_98)
```

## Again, here is an attempt to run regression with everything from economy, but only the surviving columns of listing data
```{r}
fit5 <- lm(formula = Avg..Median.Listing.Price.Yy ~ Avg..Active.Listing.Count.Yy + 
    Avg..Median.Dom + factor(Region.x) + 
End.Of.2016.Household + End.Of.2017.Household + End.Of.2016.Job + End.Of.2017.Job + Income.2017 + Income.2016 + Unemployment.Rate.2017 + Unemployment.Rate.2016 + Buy.Pct.2017 + Buy.Pct.2016 + HH_yoy + JOB_yoy + INC_yoy + UNEMP_yoy + BUYPCT_yoy + Home.Ownership.2017 + OwnOccHH2017 + Sale.Px.Recovery + Sale.Price.Yoy + Sale.Px.Recovery..Msa1. + Sale.Price.Yoy..Msa1. + Total.Starts + Total.Starts.Recovery + Total.Starts.Yoy  
 + Total.Starts..Msa1. + Total.Starts.Recovery..Msa1. + Total.Starts.Yoy..Msa1., data = listing_eco_98)

summary(fit5)
```


## Running stepwise regression:
```{r}
model.select(fit5, verbose = FALSE)

```

... we are pointed to 8 variables.

Running regression once again, with those 8 and storing the results this time
```{r}
fit6 <- lm(formula = Avg..Median.Listing.Price.Yy ~ Avg..Active.Listing.Count.Yy + 
    End.Of.2016.Household + End.Of.2017.Household + Income.2017 + 
    Unemployment.Rate.2017 + INC_yoy + BUYPCT_yoy + Sale.Price.Yoy..Msa1., 
    data = listing_eco_98)

summary(fit6)
```
```{r}
fit7 <- lm(formula = Avg..Median.Listing.Price.Yy ~ Avg..Active.Listing.Count.Yy + 
    End.Of.2016.Household + End.Of.2017.Household + Income.2017 + 
    Unemployment.Rate.2017 + INC_yoy + Sale.Price.Yoy..Msa1., 
    data = listing_eco_98)

summary(fit7)
```
```{r}
fit8 <- lm(formula = Avg..Median.Listing.Price.Yy ~ Avg..Active.Listing.Count.Yy + 
    Avg..Median.Dom + factor(Region.x) + 
End.Of.2016.Household + End.Of.2017.Household + End.Of.2016.Job + End.Of.2017.Job + Income.2017 + Income.2016 + Unemployment.Rate.2017 + Unemployment.Rate.2016 + Buy.Pct.2017 + Buy.Pct.2016 + HH_yoy + JOB_yoy + INC_yoy + UNEMP_yoy + Home.Ownership.2017 + OwnOccHH2017 + Sale.Px.Recovery + Sale.Price.Yoy + Sale.Px.Recovery..Msa1. + Sale.Price.Yoy..Msa1. + Total.Starts + Total.Starts.Recovery + Total.Starts.Yoy  
 + Total.Starts..Msa1. + Total.Starts.Recovery..Msa1. + Total.Starts.Yoy..Msa1., data = listing_eco_98)

summary(fit8)
```
```{r}
model.select(fit8, verbose = FALSE)
```

```{r}
fit9 <-lm(formula = Avg..Median.Listing.Price.Yy ~ Avg..Active.Listing.Count.Yy + 
    factor(Region.x) + End.Of.2016.Job + Income.2017 + 
    Unemployment.Rate.2017 + HH_yoy + 
    INC_yoy + Sale.Price.Yoy..Msa1., data = listing_eco_98)
summary(fit9)
```


Model from Javier's file:
----------------------------

Load data again
```{r}
listing <- read.xlsx("HES_TEMPLATES_v2.xlsx", 1)
economy <- read.xlsx("HES_TEMPLATES_v2.xlsx", sheetIndex = 2, startRow = 1, endRow = 382)

#remove labels from economic data
economy<-economy[,-c(2:5) ]

#merge
listing_eco <- merge(listing, economy, by="RankHH")

#View(listing_eco)
names(economy)
#View(economy[,5:31])

#Subset data to include largest 100 MSAs only
mydata=data.frame(listing_eco[1:100,])
View(mydata)
```


Do some clean up
```{r}
#Filter data based on initial findings

#remove labels
mydata<-mydata[,-c(1:3) ]

#remove bad explanatory variables based on known multicollinearity and data coverage issues
mydata<-mydata[,-c(6,13,14,16,18,24,27:36) ]

#remove bad rows based on known issues
mydata<-mydata[-c(1,89), ]
```


```{r}
fit=lm(Avg..Median.Listing.Price.Yy ~ 
          
          +Avg..Active.Listing.Count.Yy
       +Avg..Ldpviews.Per.Property
       + Avg.ActiveListingsper1000HH
       + HH_yoy
       +JOB_yoy
       +INC_yoy
       
       +Region
       
       , data=mydata)

summary(fit)
```

----------------------------------------

Attempt prediction on 20% sample data from the original dataset

```{r}
set.seed(121) # so that homeword and knitting of pdf are aligned
sample <- sample.int(n=nrow(mydata),size=20, replace = F)
test.mydata <- mydata[sample,]

tpredict = predict(fit,newdata=test.mydata,type="response") 
#tpredict

```

```{r}
#output<- c(tpredict, test.mydata$Avg..Median.Listing.Price.Yy)
#output
test.mydata$predicted <- tpredict
test.mydata$residual <- test.mydata$Avg..Median.Listing.Price.Yy - tpredict

test.mydata[,c("Avg..Median.Listing.Price.Yy","predicted","residual" )]


```

mean of Avg..Median.Listing.Price.Yy
```{r}
mean(test.mydata$Avg..Median.Listing.Price.Yy)
```


Mean residual:
```{r}
mean(test.mydata$residual)
```

Variance of residuals is:
```{r}
var(test.mydata$residual)
```


Let's do this 1000 times and see the results:

```{r}
repetitions=1000
means <- rep(0,repetitions)
vars <- rep(0, repetitions)
for(i in 1:repetitions){
  sample <- sample.int(n=nrow(mydata),size=20, replace = F)
  test.mydata <- mydata[sample,]
  
  tpredict = predict(fit,newdata=test.mydata,type="response") 
  tresidual = test.mydata$Avg..Median.Listing.Price.Yy - tpredict
  means[i] = mean(tresidual)
  vars[i] = var(tresidual)
  #print(i)
}

mean(means) # mean of mean of residuals
mean(vars)  # mean of variance of residuals
```










