---
title: "RE Project Notes RS"
author: "Evgeny Nenashev"
date: "April 3, 2019"
output: pdf_document
---

### Based on the initial commit of Ravi Sharma
### 
Load package for working with Excel files

```{r}
library(xlsx)
```

Load listing data

```{r}
listing <- read.xlsx("HES_TEMPLATES_v2.xlsx", 1)

dim(listing)
```

Load economy data

```{r}
economy <- read.xlsx("HES_TEMPLATES_v2.xlsx", sheetIndex = 2, startRow = 1, endRow = 382)

dim(economy)

```

Header records:

```{r}
head(listing)
```

```{r}
tail(economy)
```

## As discussed in the group (...add details), to consider a subset comprising of top 100 ranking city areas, a smaller dataset is created:
```{r}
listing100 <- data.frame(listing[1:100,])
dim(listing100)
```


## Summary of variables in Listings (top 100 ranks subset of data):
### To begin with, list all variables:
```{r}
names(listing)
summary(listing)
```

### 04. Regions represented in the data:

Summary table:
```{r}
summary(listing100$Region)
```
Visual:
```{r}
barplot(table(listing100$Region), main = "Regions in top 100 listing centers", xlab="Regions")
```

Comments:


### 05. Median.Listing.Price.Yy change:
Summary table:
```{r}
summary(listing100$Avg..Median.Listing.Price.Yy)
```
Visual:
```{r}
boxplot(listing100$Avg..Median.Listing.Price.Yy)
```


### 06. Active.Listing.Count.Yy change:
Summary table:
```{r}
summary(listing100$Avg..Active.Listing.Count.Yy)
```
Visual:
```{r}
boxplot(listing100$Avg..Active.Listing.Count.Yy)
```

### 07. Ldpviews.Per.Property.Yy change:
Summary table:
```{r}
summary(listing100$Avg..Ldpviews.Per.Property.Yy)
```
Visual:
```{r}
boxplot(listing100$Avg..Ldpviews.Per.Property.Yy)
```


### 08. Median.Dom.Yy change:
Summary table:
```{r}
summary(listing100$Avg..Median.Dom.Yy)
```
Visual:
```{r}
boxplot(listing100$Avg..Median.Dom.Yy)
```

### 09. Median.Listing.Price :
Summary table:
```{r}
summary(listing100$Avg..Median.Listing.Price)
```
Visual:
```{r}
boxplot(listing100$Avg..Median.Listing.Price)
```


### 10. Active.Listing.Count :
Summary table:
```{r}
summary(listing100$Avg..Active.Listing.Count)
```
Visual:
```{r}
boxplot(listing100$Avg..Active.Listing.Count)
```


### 11. Ldpviews.Per.Property :
Summary table:
```{r}
summary(listing100$Avg..Ldpviews.Per.Property)
```
Visual:
```{r}
boxplot(listing100$Avg..Ldpviews.Per.Property)
```


### 12. Median.Dom :
Summary table:
```{r}
summary(listing100$Avg..Median.Dom)
```
Visual:
```{r}
boxplot(listing100$Avg..Median.Dom)
```


### 13. ActiveListingsper1000HH :
Summary table:
```{r}
summary(listing100$Avg.ActiveListingsper1000HH)
```
Visual:
```{r}
boxplot(listing100$Avg.ActiveListingsper1000HH)
```




## Try regression with everything in:

#```{r}
#fit1 <- lm(Avg..Median.Listing.Price.Yy ~ Avg..Active.Listing.Count.Yy + Avg..Ldpviews.Per.Property.Yy #+ Avg..Median.Dom.Yy + Avg..Median.Listing.Price + Avg..Active.Listing.Count + #Avg..Ldpviews.Per.Property + Avg..Median.Dom + Avg.ActiveListingsper1000HH , data=listing100)
#
#summary(fit1)
#```

### Try regression with everything in:
### The same in a bit more compact way

Fit on listing data only (no economy data)

```{r}
#head(listing100)
### Removing Cbsacode and Cbsatitle
listing100_n=listing100[,c(-2,-3)]
### Keeping RankHH for subsequent merge
### NO region at all
fit1_n=lm(Avg..Median.Listing.Price.Yy~ . -Region -RankHH ,data=listing100_n)
summary(fit1_n)
```

### NO VIF above 10

```{r}
#install.packages("car")
library(car)
vif(fit1_n)
```


### Let us also include the categorical variable region to list of predictors:
### The same in a bit more compact way
```{r}
fit2_n=lm(Avg..Median.Listing.Price.Yy~ . -Region + factor(Region) -RankHH,data=listing100_n)
summary(fit2_n)
vif(fit2_n)
```


## Let us look at one-to-one relationship between Y and X and between Xs:
```{r}
pairs(listing100[,5:13])
```

## Now let us look at the suggested predictors individually, in detail:
#
#### 
#```{r}
#fit3 <- lm(Avg..Median.Listing.Price.Yy ~ Avg..Active.Listing.Count.Yy, data=listing100)
#
#summary(fit3)
#```
#Scatterplot and regression line
#```{r}
#plot(listing100$Avg..Active.Listing.Count.Yy, listing100$Avg..Median.Listing.Price.Yy)
#abline(fit3, col="orange", lwd=3)
#```
#
#The plot is somewhat convincing. Let us look for influential points
#```{r}
#plot(fit3,4)
#```
#
#Let us list the two longest on Cook's distance:
#```{r}
#listing100[c(31,89),1:4]
#```
#
#Let us remove those two:
#```{r}
#listing98<-listing100[!(listing100$RankHH == 31 | listing100$RankHH == 89),]
#
#dim(listing98)
#```

### The same with less typing
```{r}
listing98_n=listing100_n[c(-31,-89),]
dim(listing98_n)
```

#And now run the regression again:
#```{r}
#fit4 <- lm(Avg..Median.Listing.Price.Yy ~ Avg..Active.Listing.Count.Yy, data=listing98)
#
#summary(fit4)
#```
#Slightly improved Adjusted R-squared, residual standard error reduces slightly. Let us see the plots #again:
#```{r}
#plot(listing98$Avg..Active.Listing.Count.Yy, listing98$Avg..Median.Listing.Price.Yy)
#abline(fit3, col="orange", lwd=3)
#```



### Let us get economy data into the mix:
## First let us create the combined dataset:
#```{r}
#listing_eco_98 <- merge(listing98, economy, by="RankHH")
#dim(listing_eco_98)
#```
```{r}
listing_eco_98_n <- merge(listing98_n, economy, by="RankHH")
listing_eco_98_n=listing_eco_98_n[,c(-12,-13)]
dim(listing_eco_98_n)
names(listing_eco_98_n)
```

## we now have a long list of columns. Here is the list:
#```{r}
#names(listing_eco_98)
#```

## Again, here is an attempt to run regression with everything from economy, but only the surviving columns of listing data
#```{r}
#fit5 <- lm(formula = Avg..Median.Listing.Price.Yy ~ Avg..Active.Listing.Count.Yy + 
#    Avg..Median.Dom + factor(Region.x) + 
#End.Of.2016.Household + End.Of.2017.Household + End.Of.2016.Job + End.Of.2017.Job + Income.2017 + #Income.2016 + Unemployment.Rate.2017 + Unemployment.Rate.2016 + Buy.Pct.2017 + Buy.Pct.2016 + HH_yoy + #JOB_yoy + INC_yoy + UNEMP_yoy + BUYPCT_yoy + Home.Ownership.2017 + OwnOccHH2017 + Sale.Px.Recovery + #Sale.Price.Yoy + Sale.Px.Recovery..Msa1. + Sale.Price.Yoy..Msa1. + Total.Starts + Total.Starts.Recovery #+ Total.Starts.Yoy  
# + Total.Starts..Msa1. + Total.Starts.Recovery..Msa1. + Total.Starts.Yoy..Msa1., data = listing_eco_98)
#
#summary(fit5)
#```


###Lets remember we removed factor(Region.x)
### 
```{r}
head(listing_eco_98_n)
listing_eco_98_n[1,]
names(listing_eco_98_n)
fit5_n=lm(Avg..Median.Listing.Price.Yy~ . -RankHH -Region.x 
          #+factor(Region.x)
,data=listing_eco_98_n)
summary(fit5_n)
vif(fit5_n)

#pairs(listing_eco_98_n[,3:7])
#pairs(listing_eco_98_n[,8:12])
```

### Technically at this point there is no VIF higher than 10, there are some close to it though
### Removing High VIFs
#```{r}
#fit5_n=lm(Avg..Median.Listing.Price.Yy~ . -RankHH -Region.x -End.Of.2017.Household -End.Of.2017.Job #-Income.2017 -Unemployment.Rate.2017 -Buy.Pct.2017
#          -End.Of.2016.Household -End.Of.2016.Job -Sale.Px.Recovery -Sale.Px.Recovery..Msa1. #-Avg..Median.Listing.Price -Region.y -Total.Starts -Avg..Active.Listing.Count
#          -OwnOccHH2017 -Total.Starts.Recovery..Msa1.
#          ,data=listing_eco_98_n)
#summary(fit5_n)
#vif(fit5_n)
#```

### At this point no VIF greater than 10

### stepwise



```{r}
fit_n_step_forward = step(fit5_n, direction = "forward", trace=FALSE)
summary(fit_n_step_forward)
BIC(fit_n_step_forward)
```

```{r}
fit_n_step_backward = step(fit5_n, direction = "backward", trace=FALSE)
summary(fit_n_step_backward)
BIC(fit_n_step_backward)
```


```{r}
fit_n_step_both = step(fit5_n, direction = "both", trace=FALSE)
summary(fit_n_step_both)
BIC(fit_n_step_both)

```



```{r}
anova(fit_n_step_forward,fit_n_step_backward)
```


### Now removing BUYPCT_yoy


```{r}

fit17_n=lm(Avg..Median.Listing.Price.Yy~ . -RankHH -Region.x -BUYPCT_yoy 
          #+factor(Region.x)
,data=listing_eco_98_n)
#summary(fit5_n)
#vif(fit5_n)


#fit17_n=lm(Avg..Median.Listing.Price.Yy~ . -RankHH -Region.x -End.Of.2017.Household -End.Of.2017.Job #-Income.2017 -Unemployment.Rate.2017 -Buy.Pct.2017
#          -End.Of.2016.Household -End.Of.2016.Job -Sale.Px.Recovery -Sale.Px.Recovery..Msa1. #-Avg..Median.Listing.Price -Region.y -Total.Starts -Avg..Active.Listing.Count
#          -OwnOccHH2017 -Total.Starts.Recovery..Msa1.
#          -BUYPCT_yoy
#          ,data=listing_eco_98_n)
summary(fit17_n)
#vif(fit17_n)
```
#```{r}
#fit_n_step_forward = step(fit17_n, direction = "forward", trace=FALSE)
#summary(fit_n_step_forward)
#BIC(fit_n_step_forward)
#```

```{r}
fit_n_step_backward = step(fit17_n, direction = "backward", trace=FALSE)
summary(fit_n_step_backward)
BIC(fit_n_step_backward)
```

### Now lets try to remove 2017 variables since we may not have them available

```{r}
fit18_n=lm(Avg..Median.Listing.Price.Yy~ . -RankHH -Region.x -BUYPCT_yoy 
           -End.Of.2017.Household -End.Of.2017.Job -Income.2017 -Buy.Pct.2017
          #+factor(Region.x)
,data=listing_eco_98_n)
summary(fit18_n)

fit_n_step_backward = step(fit18_n, direction = "backward", trace=FALSE)
summary(fit_n_step_backward)
BIC(fit_n_step_backward)
```



